---
title: "SCTLD_BiNGO_Analysis_May2021"
author: "Melissa Drown"
date: "5/4/2021"
output: html_document
---
```{r Library Loading, include = F}
library(tidyverse)
library(DESeq2)
library(ggrepel)
library(ComplexHeatmap)
library(pheatmap)
library(genefilter)
library(WGCNA)
library(flashClust)
library(edgeR) #counts per million
library(venn)
```

This analysis uses previously generated files from differential expression analysis to make a list of annotated protein identifiers matched to GO terms for GO Enrichment analysis in Cytoscape BiNGO. After running the BiNGO analysis, results tables are used for data visualization.

## Set up Files for Gene Ontology Analysis
```{r custom annot file for bingo and cytoscape}
# Ofav reference list for BiNGO Enrichment Analysis
ofav_annot_complete$GOs <- as.character(ofav_annot_complete$GOs)

ofav_annot_complete %>% dplyr::select(Protein.Identifier, GOs) %>%
  mutate_all(na_if,"") %>% 
  drop_na() %>% 
  tidyr::separate_rows(GOs, sep = ",") %>%
  mutate(GOs = str_remove(GOs, "GO:")) %>% 
  unite(x, c(Protein.Identifier, GOs), sep = " = ", remove = TRUE) -> caf_go

View(caf_go)

#write.table(caf_go,
 #           file = "~/Desktop/cyto_ofav_input.txt",
  #          quote=FALSE, 
   #         row.names=FALSE, 
    #        col.names=FALSE)
```

```{r}
# Ofav list of FDR significant for BiNGO Enrichment Analysis
ofav_annot_DEG %>% 
  filter(padj < 0.05) %>%
  dplyr::select(Protein.Identifier) %>% 
  drop_na() %>% 
  distinct(Protein.Identifier) %>%
  column_to_rownames(var = "Protein.Identifier") %>% 
  row.names(.) -> ofav_BH

ofav_BH
length(ofav_BH) # 2005 -- 32 don't have protein_ID annotations and 157 duplicate protein_IDs

#write.table(
 # ofav_BH,
#  file = "~/Desktop/ofav_BH.txt",
 # sep = "\t",
  #quote = FALSE,
  #row.names = FALSE,
  #col.names = FALSE
#)
```

Same setup for Mcav
```{r custom annot file for bingo and cytoscape}
# Mcav Reference list for BiNGO Enrichment Analysis
mcav_annot_complete$GOs <- as.character(mcav_annot_complete$GOs)

mcav_annot_complete %>% dplyr::select(prot_id, GOs) %>%
  mutate_all(na_if,"") %>% 
  drop_na() %>% 
  tidyr::separate_rows(GOs, sep = ",") %>%
  mutate(GOs = str_remove(GOs, "GO:")) %>% 
  unite(x, c(prot_id, GOs), sep = " = ", remove = TRUE) -> maf_go

View(maf_go)

#write.table(maf_go,
 #           file = "~/Desktop/cyto_mcav_input.txt",
  #          quote=FALSE, 
   #         row.names=FALSE, 
    #        col.names=FALSE)
```

```{r}
# Mcav list of FDR significant for BiNGO Enrichment Analysis
mcav_annot_DEG %>% 
  filter(padj < 0.05) %>%
  dplyr::select(prot_id) %>% 
  drop_na() %>% 
  distinct(prot_id) %>%
  column_to_rownames(var = "prot_id") %>% 
  row.names(.) -> mcav_BH

mcav_BH
length(mcav_BH) # 582


#write.table(
#  mcav_BH,
#  file = "~/Desktop/mcav_BH.txt",
#  sep = "\t",
#  quote = FALSE,
#  row.names = FALSE,
#  col.names = FALSE
#)
```

Visualizing the GO results
```{r go viz}
ofav_go_bp <- read.csv("~/Desktop/ofav_BiNGO_signifGO_BP.csv")

ofav_go_cc <- read.csv("~/Desktop/ofav_BiNGO_signifGO_CC.csv")

# Biological Process Ofav BiNGO Results
ofav_go_bp$GO_description <- as.character(ofav_go_bp$GO_description)
ofav_go_bp$genes <- as.character(ofav_go_bp$genes)
ofav_go_bp$GO_ID <- as.factor(ofav_go_bp$GO_ID)
ofav_go_bp$protein_product <- as.character(ofav_go_bp$genes)

# Molecular Function Ofav BiNGO Results
ofav_go_cc$GO_description <- as.character(ofav_go_cc$GO_description)
ofav_go_cc$genes <- as.character(ofav_go_cc$genes)
ofav_go_cc$GO_ID <- as.factor(ofav_go_cc$GO_ID)
ofav_go_cc$protein_product <- as.character(ofav_go_cc$genes)
```

```{r}
#Function for Formatting Bingo Heatmaps
pad27 <- function(x){
  str_x = as.character(x)
  digits = nchar(str_x)
  toAdd = 7-digits
  name = paste0('GO:',
                strrep(0,toAdd), 
                x)
  name
}

ofav_annot_DEG$protein_product <- ofav_annot_DEG$Protein.Identifier

# seperate list of protein_products associated with each GO (one protein product per line, for merging with DEG list)
ofav_go_bp %>% 
  separate_rows(protein_product, sep = '\\ ', convert = TRUE) %>%
   mutate(GO.ID_FULL = pad27(GO_ID)) %>%
  inner_join(ofav_annot_DEG, by="protein_product") -> ofav_BP


ofav_go_cc %>% 
  separate_rows(protein_product, sep = '\\ |', convert = TRUE) %>% 
  mutate(GO.ID_FULL = pad27(GO_ID)) %>%
  inner_join(ofav_annot_DEG, by="protein_product") -> ofav_CC
```

```{r}
# plot p-value and number of genes per GO term for Ofav BP and CC list
# no significant GO terms for Mcav for any category so there is nothing to plot
#library(scales)
#library(viridis)

ggplot(data=ofav_go_bp, aes(x=-log(padj), y=reorder(GO_description, -padj))) +
    geom_point(aes(size=number_genes, color=-log(padj))) +
    scale_color_viridis(option="plasma") +
  labs(x="-log FDR Adjusted P-Value", y="Gene Ontology Term Description", title=paste("Ofav Significant Biological Process GO")) +
  scale_y_discrete(labels = wrap_format(40)) +
  theme_bw()

ggplot(data=ofav_go_cc, aes(x=-log(padj), y=reorder(GO_description, -padj))) +
    geom_point(aes(size=number_genes, color=-log(padj))) +
    scale_color_viridis(option="plasma") +
  labs(x="-log FDR Adjusted P-Value", y="Gene Ontology Term Description", title=paste("Ofav Significant Cellular Component GO")) +
  scale_y_discrete(labels = wrap_format(40)) +
  theme_bw()

```
